---
title: "Oaxaca Isoscape Data Preparation"
author: Nick Gauthier
format: html
editor: visual
---

## Setup

Load required packages.

```{r setup}
# data management, processing, visualization
library(tidyverse) # for data manipulation
library(readxl) # for reading Excel file
library(stars) # for raster data
library(terra) # for raster data
library(sf) # for spatial data
library(rnaturalearth) # for world map data
library(patchwork)
library(here)
library(dggridR)
source(here('R/helpers.R'))
```

## Preprocess Global Raster Predictors

### Bedrock isotope predictors

Get the list of predictor files and names produced from the original analysis by Bataille et al. (2020).

```{r}
pred_files <- '../data/raw/isotope_predictors' |>
  list.files(pattern = '\\.tif$', full.names = TRUE)

pred_names <- pred_files |>
  str_remove('../data/raw/isotope_predictors/') |>
  str_remove('\\.tif$') |>
  str_remove('_reproj') |>
  str_remove('r\\.')

sediment_tibble <- tibble(
  id = c(0, 1, 2, 3, 4, 5, 6, 7, 8, 9, 10),
  category = c('none', 'colluvial', 'eolian', 'glacial', 'lacustrine', 
               'marine', 'organic', 'evaporite', 'pyroclastics', 'coastal', 'ice')
)
```

Some of the variables have slightly different origins and extents, so we need to process them step by step before combining them.

```{r}
vars1 <- rast(pred_files[c(4, 14, 15, 18)])
names(vars1) <- pred_names[c(4, 14, 15, 18)]
levels(vars1[[2]]) <- sediment_tibble
names(vars1[[2]]) <- 'sediment'

vars2 <- rast(pred_files[c(1,2,3,20,21)])
crs(vars2) <- crs(vars1)
names(vars2) <- pred_names[c(1,2,3,20,21)]

predictors <- c(vars1, vars2)
bedrock_predictors <- predictors[[4:9]]
other_predictors <- predictors[[1:3]]

bedrock_pcs <- prcomp(bedrock_predictors, scale. = TRUE, maxcell = 10000000, retx = FALSE)
bedrock_pc_names <- c("bedrock_age_sr_gradient", "bedrock_young_highSrVar"#, "bedrock_ageRange_SrContrast"
                      )
```

### Soil

Soilgrids 250m, already aggregated to 1km.

```{r}
soils <- here('data/raw/soilgrids') |>
  list.files(full.names = TRUE) |>
  rast()

names(soils) <- names(soils) |>
  str_replace('-', '_')

soil_pcs <- prcomp(soils, scale. = TRUE, maxcell = 10000000, retx = FALSE)

soil_pc_names <- c(
  "soil_moisture_and_bulk_density",  # PC1: High phh2o and moderate bulk density; contrast with clay
  "soil_clay_cec_gradient",          # PC2: Contrast between higher clay and lower cec
  "soil_clay_cec_texture"#,           # PC3: Combined effect of clay and cec (texture), not limited to surface layers
#  "soil_depth_dependent_bulk_density", # PC4: Reflects vertical differences in bulk density
 # "soil_chemical_mineral"#,           # PC5: Captures additional chemical/mineral variability
#  "soil_soc_enrichment",             # PC6: Dominated by high organic carbon (SOC) enrichment
 # "soil_soc_contrast"                # PC7: Inverse SOC pattern relative to PC6, with additional clay/cec influence
)
```

### Elevation
```{r}
elevation <- geodata::elevation_global(0.5, path = here('data/raw'))
names(elevation) <- 'elevation'
```

### Climate
```{r}
# climate data
climate <- list.files(here('data/raw/CHELSA_V2_bio_clim'), full.names = TRUE) |>
  rast()

names(climate) <- names(climate) |>
  str_remove('CHELSA_') |>
  str_remove('_1981-2010_V.2.1') |>
  str_remove('1981-2010_') |>
  str_remove('_V.2.1')

climate_pcs <- prcomp(climate, scale. = TRUE, maxcell = 10000000, retx = FALSE)

climate_pc_names <- c(
  "climate_temp_dryness_gradient",      # PC1: Captures the overall temperature–dryness gradient.
  "climate_moisture_precipitation",       # PC2: Represents the moisture and precipitation regime.
  "climate_temperature_extremes",         # PC3: Reflects temperature extremes and seasonality.
  "climate_seasonal_precipitation",       # PC4: Indicates seasonal precipitation patterns.
  "climate_precipitation_intensity_timing", # PC5: Describes precipitation intensity and timing.
  "climate_radiation_vs_wind_dryness",    # PC6: Contrasts radiation with wind and dryness.
  "climate_subtle_moisture_regime"        # PC7: Captures subtle aspects of the moisture regime.
)
```

### Aerosol Deposition
Dry and wet aerosol deposition from MERRA-2. Run `R/aerosol_deposition.R` first to get the data ready.
```{r}
aerosols_stars <- readRDS(here('data/derived/aerosols.rds'))
  
aerosols <- rast(aerosols_stars)
names(aerosols) <- names(aerosols_stars)

aerosol_pcs <- prcomp(aerosols, scale. = TRUE, maxcell = 10000000, retx = FALSE)

aerosol_pc_names <- c(
  "aerosol_carbonaceous_sulfate_abundance",     # Overall loading of carbon and sulfate species, representing total anthropogenic pollution
  "aerosol_sea_salt_signature",                 # Distinguishes marine aerosols (sea salt) from other components
  "aerosol_dust_concentration",                 # Represents mineral dust loading, likely from arid regions or long-range transport
  "aerosol_dry_wet_sulfate_organic_contrast",   # Contrasts dry sulfate with wet organic carbon, indicating different processing pathways
  "aerosol_dry_carbon_dominance"                # Highlights the importance of dry carbonaceous aerosols vs. wet components
)
```

### Pyroclastic Deposits
Get the extra pyroclastic layer from GUM
```{r}
sf_use_s2(FALSE)
gum_pyroclastic <- read_sf('~/Downloads/Boerker_et_al_GUM_v1.0/GUM_pyroclastics/GUM_pyroclastics.shp') |>
  st_union() |>
  st_as_sf() |>
  st_transform(eck4)
sf_use_s2(FALSE)
```

### Fertilizer application
```{r}
fertilizer <- rast(c('~/Downloads/NFertilizer_global_geotif/nfertilizer_global.tif',
              '~/Downloads/PFertilizer_global_geotif/pfertilizer_global.tif')) |>
  setNames(c('nitrogen', 'phosphorus')) |>
  subst(NA, 0)
```


## Mexico Predictor Data

This is too much data to keep in memory. Because we're only making spatial predictions for Mexico, we'll crop the predictors to the bounding box of Mexico so we can work with a smaller dataset.

Define a function to fill gaps in the rasters.
```{r}
fillNA <- function(x, method = 'mean') {
  focal(x, w = 3, fun = method, na.policy = 'only', na.rm = TRUE) |>
    focal(w = 3, fun = method, na.policy = 'only', na.rm = TRUE) |>
    focal(w = 3, fun = method, na.policy = 'only', na.rm = TRUE) |>
    focal(w = 3, fun = method, na.policy = 'only', na.rm = TRUE) |>
    focal(w = 3, fun = method, na.policy = 'only', na.rm = TRUE) |>
    focal(w = 3, fun = method, na.policy = 'only', na.rm = TRUE) |>
    focal(w = 3, fun = method, na.policy = 'only', na.rm = TRUE) |>
    focal(w = 3, fun = method, na.policy = 'only', na.rm = TRUE)
}
```

```{r}
mexico_bbox <- st_bbox(c(xmin = -106, xmax = -87, ymin = 15, ymax = 23), crs = 4326) |>
  st_transform(eck4) 

mexico_bedrock <- crop(bedrock_predictors, mexico_bbox) |>
  fillNA(method = 'modal') |>
  predict(bedrock_pcs, index = 1:2, cores = 4)
names(mexico_bedrock) <- bedrock_pc_names

mexico_other <- crop(other_predictors, mexico_bbox) |>
  fillNA(method = 'modal')

mexico_climate <- crop(climate, st_transform(mexico_bbox, 4326)) |> 
  predict(climate_pcs, index = 1:7, cores = 4) |>
  project(mexico_other)
names(mexico_climate) <- climate_pc_names

mexico_soils <- crop(soils, st_transform(mexico_bbox, crs(soils))) |> 
  predict(soil_pcs, index = 1:3, cores = 4) |>
  project(mexico_other) |>
  fillNA('mean')
names(mexico_soils) <- soil_pc_names

mexico_aerosols <- crop(aerosols, st_transform(mexico_bbox, 4326), snap = 'out') |> 
  predict(aerosol_pcs, index = 1:5, cores = 4) |>
  project(mexico_other, method = 'cubicspline')
names(mexico_aerosols) <- aerosol_pc_names

mexico_fertilizer <- crop(fertilizer, st_transform(mexico_bbox, 4326), snap = 'out') |>
  project(mexico_other, method = 'cubicspline')

mexico_elevation <- crop(elevation, st_transform(mexico_bbox, 4326)) |> 
  project(mexico_other)

template <- is.na(mexico_climate[["climate_temp_dryness_gradient"]]) |>
  mask(mexico_climate[["climate_temp_dryness_gradient"]]) |>
  st_as_stars()

mexico_pyroclastic <- gum_pyroclastic |>
  st_rasterize(template) |>
  setNames('pyroclastic') |>
  rast()

mexico_predictors <- c(mexico_bedrock, mexico_other, mexico_pyroclastic, mexico_aerosols, mexico_climate, mexico_soils, mexico_elevation, mexico_fertilizer) %>%
  mask(., .[[15]])

names(mexico_predictors)
```


```{r}
plot(mexico_bedrock)
plot(mexico_aerosols)
plot(mexico_climate)
plot(mexico_soils)
plot(mexico_other)
plot(mexico_fertilizer)
```

```{r}
plot(mexico_predictors[[1:12]])
plot(mexico_predictors[[13:24]])
```

Save the Mexico predictors to disk so we don't have to reprocess them each time.

```{r}
saveRDS(mexico_predictors, here('data/derived/mexico_predictors.rds'))
writeRaster(mexico_predictors, here('data/derived/mexico_predictors.tif'), overwrite = TRUE)
mexico_predictors <- readRDS(here('data/derived/mexico_predictors.rds'))
```

## Global Isotope Observations

Now we'll load the global isotope observations from Bataille et al. (2020) and filter them to only include the types of samples we're interested in (animals, plants, soil, and water) and remove any samples with Sr isotope ratios greater than 0.75 (to remove any outliers that might be due to contamination or other issues). We'll also convert the data to a spatial object and transform it to the Eckert IV projection.

```{r}
pts1 <- here('data/raw/facets-2024-0180supplb.xlsx') |>
  read_excel(sheet = 2, guess_max = 1000000) |>
  st_as_sf(coords = c('Longitude', 'Latitude'), crs = 4326) |>
  st_transform(eck4) |>
  select(Sr = `87Sr/86Sr`, type = Type_1, type2 = Type2, material = Material) |>
  mutate(type = tolower(type), type2 = tolower(type2), material = tolower(material)) |>
  filter(type != 'food',
         type != 'insect') |>
  mutate(type = if_else(type %in% c('soil', 'minerals'), 'soil', type),
         type = if_else(type %in% c('mammals', 'mammal', 'reptile', 'invertebrate', 'amphibian', 'biomineral'), 'animal', type)) 

not_near <- function(x, y, dist) !st_is_within_distance(x, y, dist)
pts2 <- here('data/raw/Wang et al./data/Wang_et_al_Dataset_S1.xlsx') |>
  read_excel(skip = 2, guess_max = 1000000) |>
  st_as_sf(coords = c('Longitude', 'Latitude'), crs = 4326) |>
  st_transform(eck4) |>
  st_filter(st_union(pts1), .predicate = not_near, dist = 500) |>
  select(Sr = `87Sr/86Sr`, material = `Material Type`, type2 = `Taxa of plants and animals`) |>
  mutate(material = tolower(material), type2 = tolower(type2),
         type = if_else(material == 'plant', 'plant', if_else(material == 'Soil leachates', 'soil', 'animal'))) |>
  filter(material != 'insect') 

pts3 <- here('data/raw/1-s2.0-S0031018220302947-mmc1.xlsx') |>
  read_excel(sheet = 2, guess_max = 1000000) |>
  filter(Source.of.data.reference.ID != 144, # location centroids not correct
         Source.of.data.reference.ID != 45) |>  # extreme variation in seasonal water at one site
  filter(`Geolocation Uncertainty` != 'Very High' | is.na(`Geolocation Uncertainty`)) |>
  mutate(Latitude = if_else(Country %in% c('Peru', 'Chile') & Latitude > 0, Latitude * -1, Latitude)) |> # some human samples have the wrong hemisphere
  select(Sr = `87Sr/86Sr`, type = `Type 1`, type2 = `Type 2`, type3 = `Type 3`, 
         material = Material, Longitude, Latitude) |>
  filter(!is.na(Sr)) |>
  mutate(across(type:material, tolower),
         Sr = str_replace(Sr, ',', '.'),
         Sr = parse_number(Sr)) |>
  st_as_sf(coords = c('Longitude', 'Latitude'), crs = 4326) |>
  st_transform(eck4) |>
  filter(!(type %in% c('bird', 'insect', 'dust'))) |>
  mutate(type2 = if_else(is.na(type2) & type3 %in% c('snail', 'snail shell'), 'snail', type2),
         type2 = if_else(is.na(type2), type, type2),
         type2 = if_else(type2 == "–", type, type2),
         type = if_else(type2 == 'human', 'human', type)) |>
  select(-type3)

water <- pts3 |>
  filter(type == 'water',
         !(type2 %in% c('snow', 'tap', 'rain', 'throughfall', 'mineral', 'minewater', 'sewage')))

cambio <- read_excel(here('data/raw/CAMBIO Humans Database 2024 - Local Sr Only.xlsx'),
                     sheet = 1,  guess_max = 10000) |>
  st_as_sf(coords = c('Longitude', 'Latitude'), crs = 4326) |>
  st_transform(eck4) |>
  select(Sr = `87Sr/86Sr`, type2 = Culture) |>
  mutate(type = 'human')

humans <- pts3 |> # need to double check the human samples that get in here.
  filter(type == 'human') |>
  st_filter(st_union(cambio), .predicate = not_near, dist = 2) |>
  bind_rows(cambio) 

pts <- bind_rows(pts1, pts2, water, humans) |>
  mutate(oaxaca = FALSE)
```

```{r}
library(mapview)
mapview(pts3 |> st_jitter(500) |> filter(Source.of.data.reference.ID == 45), zcol = 'Sr')
mapview(pts1, color = 'yellow') + mapview(pts2, color = 'green') + mapview(pts3 |> filter(type == 'human') |> st_jitter(500), color = 'orange') + mapview(cambio |> st_jitter(500), color = 'red') + mapview(pts35, color = 'purple')
```

#### Preprocess Oaxaca Isotope Observations

Now we'll load the Oaxaca isotope observations and filter them to only include the types of samples we're interested in (plants) and remove any samples with Sr isotope ratios greater than 0.78 (to remove any outliers that might be due to contamination or other issues). We'll also convert the data to a spatial object and transform it to the Eckert IV projection.

```{r}
oaxaca_plants <- read_excel(here('data/raw/OAX Sr Result Table.xlsx')) %>%
  st_as_sf(coords = c('UTM-E', 'UTM-N'), crs = 32614) |>
  st_transform(eck4) |>
  rename(elevation = `Elevation (masl)`,
         rooting_depth = `Plant Rooting Depth`) |>
  #filter(Locality != 'San Martin Tilcajete')
  filter(Sr < 0.71)

plot(oaxaca_plants)
ggplot(oaxaca_plants, aes(Locality, Sr)) +
  geom_point()

mapview::mapview(oaxaca_plants |> st_jitter(500), zcol = 'Sr')

pts_combined <- oaxaca_plants |>
  select(Sr) |>
  mutate(type = 'plant',
         oaxaca = TRUE) |>
  bind_rows(pts) |>
  filter(between(Sr, .703, .78))
```

```{r}
saveRDS(pts_combined, here('data/derived/pts_combined.rds'))
```

```{r}
pts_combined <- readRDS(here('data/derived/pts_combined.rds'))
```

```{r}
mapview::mapview(pts_combined, zcol = 'type')
```


## Final Data for Modeling

```{r}
elevation_extract <- terra::extract(elevation, 
                                    st_transform(pts_combined, 4326), 
                                    ID = FALSE, 
                                    search_radius = 5000,
                                    #  cell = TRUE,
                                    na.rm = TRUE) |>
  select(elevation)#, cell)

extract_fun <- function(x, rast, crs = 4326){
  terra::extract(rast[[x]], st_transform(pts_combined, crs = crs), 
                 ID = FALSE, 
                 search_radius = 5000,
                 na.rm = TRUE)[,1, drop = FALSE]
}  

climate_extract <- map_dfc(1:nlyr(climate), extract_fun, climate) %>%
  predict(climate_pcs, .) %>%
  .[,1:7] %>%
  as_tibble() |> 
  setNames(climate_pc_names)

soils_extract <- map_dfc(1:nlyr(soils), extract_fun, soils, crs(soils)) %>%
  predict(soil_pcs, .) %>%
  .[,1:3] |>
  as_tibble() |> 
  setNames(soil_pc_names)

aerosol_extract <- terra::extract(aerosols, 
                                  st_transform(pts_combined, 4326), 
                                  ID = FALSE, 
                                  method = 'bicubic', 
                                  na.rm = TRUE) %>%
  predict(aerosol_pcs, .) %>%
  .[,1:5] |>
  as_tibble() |>
  setNames(aerosol_pc_names)


bedrock_extract <- map_dfc(1:nlyr(bedrock_predictors), extract_fun, bedrock_predictors, eck4) %>%
  predict(bedrock_pcs, .) %>%
  .[,1:2] |>
  as_tibble() |>
  setNames(bedrock_pc_names)

other_predictors[['sediment']]  <- as.int(other_predictors[['sediment']]) #dunno why this is necessary!
other_extract <-  map_dfc(1:nlyr(other_predictors), extract_fun, other_predictors, eck4) 

fertilizer_extract <- terra::extract(fertilizer, 
                                  st_transform(pts_combined, 4326), 
                                  ID = FALSE, 
                                  method = 'bicubic', 
                                  na.rm = TRUE)
```

```{r}
plot(other_predictors['sediment'] |> crop(st_bbox(oaxaca_plants)))
```


```{r}
mapview::mapview(dat, zcol = 'sediment')
plot(mexico_predictors['sediment'])
mapview::mapview(mexico_predictors['sediment']) + mapview::mapview(dat, zcol = 'sediment')

mapview::mapview(mexico_other['basement_age'])

plot()
```

```{r}
dat <- pts_combined |> 
  bind_cols(bedrock_extract) |>
  bind_cols(climate_extract) |>
  bind_cols(aerosol_extract) |>
  bind_cols(elevation_extract) |> 
  bind_cols(other_extract) |>
  bind_cols(soils_extract) |>
  bind_cols(fertilizer_extract) |>
  mutate(pyroclastic = st_intersects(pts_combined, gum_pyroclastic, sparse = FALSE)[,1]
) |>
  mutate(type = factor(type)) |>
  mutate(elevation = as.numeric(elevation)) |>
  mutate(sediment = factor(sediment, labels = sediment_tibble$category[1:10])) |>
  select(-geometry) # hack to move to last column
  # group_by(cell) %>%
  # mutate(
  #   n_points = n(),
  #   median_sr = median(Sr, na.rm = TRUE),
  #   mad_sr = median(abs(Sr - median_sr), na.rm = TRUE),
  #   # Only calculate thresholds and flag outliers if enough points
  #   threshold = if_else(n_points >= 5,
  #     case_when(
  #       median_sr < 0.710 ~ min(3 * mad_sr, 0.0008),
  #       TRUE ~ min(3 * mad_sr, 0.0036)
  #     ),
  #     NA_real_
  #   ),
  #   is_outlier = if_else(!is.na(threshold), abs(Sr - median_sr) > threshold, FALSE)
  # ) %>%
  # ungroup() |>
  # select(-c(n_points, median_sr, mad_sr, threshold, cell))
#select(-cell)
```

### Spatial binning

Use a discrete global grid to create equal area spatial bins for the global data

```{r}
dggs <- dgconstruct(res = 4)
grid <- dgearthgrid(dggs) |>
  st_wrap_dateline() |>
  st_transform(eck4)
plot(grid)

coords <- dat |>
  st_transform(4326) |>
  st_coordinates()

dat$seqnum <- dgGEO_to_SEQNUM(dggs, coords[,1], coords[,2])$seqnum

ggplot() +
  geom_sf(data = grid) +
  geom_sf(data = dat, aes(color = seqnum))
```

Save the final result.
```{r}
saveRDS(dat, here('data/derived/dat.rds'))
```


### Visualizations
Visualize the results. What's the distribution of global Sr values? What's the distribution of sample types?

```{r, fig.width=10, fig.height=8}
world <- ne_countries() |>
  st_transform(eck4)

a <- pts_combined |>
  arrange(Sr) |>
ggplot() +
  geom_sf(data = world, color = 'lightgrey') +
  geom_sf(aes(color = Sr), size = 0.1) +
  scale_color_viridis_c() +
  theme_minimal()

b <- ggplot(pts_combined) +
  geom_sf(data = st_as_sfc(world), color = 'lightgrey') +
  geom_sf( alpha = 0.5, size = .1) +
  theme_bw() +
  facet_wrap(~type)

c <- ggplot(pts_combined) +
  geom_histogram(aes(Sr)) +
  theme_bw() +
  labs(y = 'Count')

d <- ggplot(pts_combined) +
  geom_bar(aes(reorder(type, type, function(x) length(x)))) +
  theme_bw() +
  coord_flip() +
  labs(x = '', y = 'Count')

a / (b + (c / d) + plot_layout(widths = c(2,.7))) + plot_annotation(tag_levels = 'A', tag_suffix = ')') + 
  plot_layout(heights = c(2, 1.5))
#ggsave('data.png', width = 10, height = 8)
```

```{r fig.width = 8, fig.asp = 1}
mexico <- st_crop(world, mexico_bbox)
e <- st_crop(pts_combined, mexico_bbox) |>
  arrange(Sr) |>
  ggplot() +
  geom_sf(data = mexico, color = 'lightgrey') +
  # geom_sf(data = iso, color = 'red') +
  geom_sf(aes(color = Sr)) +
  scale_color_viridis_c() +
  theme_minimal()

f <- ggplot(st_crop(pts_combined, mexico_bbox)) +
  geom_sf(data = st_as_sfc(mexico), color = 'lightgrey') +
  geom_sf( alpha = 0.5) +
  theme_bw() +
  facet_wrap(~type)

g <- ggplot(st_crop(pts_combined, mexico_bbox)) +
  geom_histogram(aes(Sr)) +
  theme_bw() +
  labs(y = 'Count')

h <- ggplot(st_crop(pts_combined, mexico_bbox)) +
  geom_bar(aes(reorder(type, type, function(x) length(x)))) +
  theme_bw() +
  coord_flip() +
  labs(x = '', y = 'Count')

e / (f + (g / h) + plot_layout(widths = c(2,.7))) + plot_annotation(tag_levels = 'A', tag_suffix = ')') + 
  plot_layout(heights = c(2, 1.5))

#ggsave('data_mexico.png', width = 8, height = 8)
```

